### 2024-08-01 Paper/paper2.tex
Updates from Overleaf

    \begin{abstract}
		{\small
         Firm leaders are shown to be correlated with firm decisions and performance. In this paper, I investigate this topic in the context of hospital executive teams. Specifically, I empirically the effect of clinically trained executives on hospital behavior in response to an incentive shock. In particular, I leverage policies changing quality incentives for hospitals in the US in 2012, and estimate the difference in readmission and mortality rate responses for nonprofit hospitals with and without clinically trained executives. I find that a lack of clinical executives leads hospitals to improve quality more drastically after the incentive change. Further, I provide evidence that the more drastic response is due to non-clinical executives being more profit driven, and that clinical executives are not just a signal of underlying hospital objectives, but manage the hospital differently. 
		} 
	\end{abstract}

@@

    Across firm types and industries, firm leader characteristics such as gender, age, and background have been shown to be correlated with behavior and performance. I investigate this topic in the context of hospital executives with differing backgrounds: some with training as a medical doctor and some without. Physicians bring a unique combination of clinical and administration expertise, and they have insights that can potentially improve hospital outcomes (\cite{Stajduhar_2023}, \cite{Ahmed_2022}). However, hiring physician executives can also be controversial, as physicians may not always be adequately trained to enter into high level management positions (\cite{HarvardBusinessReview2018}).

    In this paper, I investigate how having clinically trained executives in leadership affects quality of care in hospitals. I scrape executive team names and titles from publicly available Tax Form 990s, filled out by nonprofit firms in the US. Combining this with publicly available hospital characteristics, I construct a novel data set of nonprofit hospitals from 2010-2014 containing information on both executive team characteristics and quality metrics. Nonprofit behavior in the hospital industry directly affects health care consumers, as private nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds. For comparison, for-profits make up 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}).

    The non-random selection of leaders makes it difficult to interpret any direct comparison of differing executive teams as causal. Therefore, I leverage an exogenous shock to hospital incentives, pay-for-performance policies enacted in the US in 2012, and compare the change in readmission and mortality rates of executive teams with and without clinical executives. Under the assumption that the two types of teams would have had similar trends in readmission and mortality rates absent the incentive change, I find that a lack of clinically trained executives leads a hospital to respond more drastically to pay-for-performance incentives. That is, while both types of hospital leadership teams decrease readmission rates after the policy change, those without clinical executives decrease readmissions at a faster rate than those with clinical executives.

    Having found that hospital leadership does, in fact, matter to hospital behavior, I investigate several mechanisms that might contribute to this difference. First, the more drastic decrease in readmissions by nonclinical teams is not driven by decreasing average patient complexity. Further, comparing each type of leadership team with for-profit hospitals reveals that the absence of clinical executives leads a nonprofit hospital to respond more similarly to a for-profit hospital. That is, having clinical training on a leadership team leads a hospital to act less profit-driven than their general training counterparts. Finally, I show that changes in a hospital's propensity to hire a clinical executive are not correlated with the pay-for-performance policies. Thus, I exploit such changes in leadership teams to decompose the effect of clinical executives into signaling (clinical executives signal underlying hospital preferences), or managing (clinical executives manage the hospital differently). I find that the entire difference in response is driven by a managing effect.

    Many researchers have documented a correlation between executive characteristics, specifically for CEOs, and firm performance, starting with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who show that manager fixed effects are an important driver of many firm behaviors and decisions. Female board members are associated with better oversight and more women executives (\cite{matsa2011chipping}; \cite{adams2009women}), but are negatively correlated with firm value (\cite{ahern2012changing}). Having more female executives is correlated with female employee wages and corporate strategies (\cite{flabbi2019female}; \cite{matsa2013female}); young male CEOs tend to be more aggressive in mergers and acquisitions, while those with military experience are less aggressive (\cite{levi2010deal}; \cite{benmelech2015military}); CEOs with general ability tend to receive higher pay and perform better (\cite{kaplan2012ceo}; \cite{custodio2013generalists}; \cite{adams2018director}; \cite{frydman2019rising}). Further, Chief Diversity Officers are not found to have any effect on hiring more diversely in universities (\cite{bradley2022impact}).

    Due to data limitations, many of these studies focus on large, publicly traded firms in the US. \citeauthor{brickley2010board} (\citeyear{brickley2010board}) have the only study to my knowledge answering this question in the US nonprofit context. Using exogenous variation in expected Medicare profits, the authors find that having an internal board of directors increases CEO compensation, and having physician board members decreases public donations (\cite{brickley2010board}). Additionally, two studies focus on hospital performance in other countries. \citeauthor{janke2019impact} (\citeyear{janke2019impact}) use data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}). However, \citeauthor{otero2022managers} (\citeyear{otero2022managers}) investigate the role of CEOs in public hospitals in Chile, and finds an 8\% decrease in mortality rates for hospitals with top managers. Further, a major contributor of this quality improvement is the exodus of older physicians as CEOs (\cite{otero2022managers}). I contribute to our knowledge of how executive team characteristics affect firm behaviors by (1) considering the important setting of nonprofit hospitals, which serve the majority of patients in the US, (2) considering a unique executive characteristic which is plausibly relevant, and (3) observing firm behavior in the presence of an exogenous incentive change.

@@

    Nonprofit firms in the US differ from for-profits mainly in that for-profit firms distribute revenue to stakeholders, while nonprofits reinvest profit back into the firm. Hospitals are governed by a board of directors, whose role is to set broad goals and strategies, and provide general oversight. The board selects executives, the highest level of management of a firm, to carry out day-to-day operations. An executive team usually consists of at least a Chief Executive Officer (CEO) and Chief Financial Officer (CFO), but there is variation in how firms organize executive teams. Some hospital executives to specialize in health care settings by earning a degree in health care management, or an MBA specific to health care. There are executive positions that are often filled by someone with clinical training, such as a Chief Medical Officer (CMO), but medical doctors can pursue other top managerial positions as well such as CEO or president. While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.

    Hospital executive teams are understudied partly due to the inaccessibility of granular data. I construct a novel data set of not-for-profit hospitals in the US that contains names, titles, and positions of all board members and executives tied to the hospital from 2009-2015. I gather this data from publicly available Tax Form 990s, which every sufficiently large nonprofit files each year in the US. To my knowledge, this is the first large scale gathering of executive names from these forms.\footnote{\citeauthor{brickley2010board} (\citeyear{brickley2010board}) collects compensation data for a select number of hospitals.}

@@

    The historical tax forms for all not-for-profits are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the NonProfit Explorer API to extract Employee Identification Numbers (EIN) of NFPs with a hospital National Taxonomy of Exempt Entities (NTEE) code. After filtering out associations and specialty hospitals, there are approximately 3,000 NFP EINs. Linking these EINs to other hospital information is crucial for the analysis. I therefore rely on matching by location and name of hospital in both the tax forms and AHA data. In the AHA data, I limit to general hospitals. Then, I confidently match 1200 EINs to an AHA identification number based on exact name matches within the same state. I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.

    For each of the matched EINs, I extract web URLs to Form 990 PDFs for each year. I use an algorithm to download these locally and use optical character recognition (OCR) text extraction methods to scrape the relevant section of the PDF, which is titled ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees". Finally, I use string cleaning methods to identify all names and positions under each EIN, year. With each name, I extract titles indicating clinical training: MD, Dr., or DO. Using OCR extraction can be unreliable in cases where the text on the form is handwritten or small. Therefore, for observations that are missing after the initial extraction, I manually record names, positions, and titles. As I limit the sample to hospitals who have text information in every year, I limit the years to 2010-2014 to retain more hospitals. This yields 852 not-for-profit hospitals that are matched to AHA data, and with complete text information in each year.

    Finally, I match each name with the National Plan and Provider Enumeration System (NPPES) database of all registered physicians. The goal is to match each physician with an NPI number in order to merge publicly available data on physician specialty, gender, and tenure. For any person who has a doctor title and uniquely matches the NPPES data, I record their NPI number. For anyone who has a doctor title and no matches, no doctor title and matches, or a doctor title and multiple matches, I use information on location and hospital name to manually match the NPI. I present a table of means for various characteristics of doctor executives in the data in Table \ref{doc_sumstats}. In total, there are 813 doctor executives. On average, they are 52 years old, and 11\% of them are female. While it is more likely they are a Chief Medical Officer than a Chief Executive Officer, still less than a third are CMOs. The most common specialty is internal medicine.

@@

    For the purposes of this paper, I focus on executive teams as a whole instead of specific positions. The main reason for this is because the organization of teams varies so drastically across hospitals that a specific position, even if it exists in two different hospitals, can be meaningfully different. I drop anyone who is solely a board member, and focus on those who have executive in their title, or are labeled as a president or vice president of a department. I create various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer. For my main sample, I limit to hospitals who do not change their propensity to hire a clinical executive for the duration of the sample.

@@

    Therefore, I focus on outcomes of readmission and mortality rates, which are publicly available at the hospital-condition level in CMS Hospital Compare. I combine rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions) into a weighted average based on the number of patients in each condition. Additionally, I include several relevant hospital characteristics from the AHA survey: number of beds, and whether the hospital is an academic medical center, is physician owned, or is system affiliated. Additionally, case mix index, a measure of overall patient complexity, comes from the CMS Case Mix Index files. Whether the hospital was penalized under HRRP or received payments under HVBP is found in the HCRIS data.

@@

    In Table \ref{tab:sumstats_samples_stable}, I present a table of means of relevant variables for several hospital samples used throughout the analysis. First, in column (1), are means for hospitals who ever have a clinical executive at any point in the sample. In column (2) are means for hospitals who have a clinical executive for the entire sample period, and column (3) shows means for hospitals who never have a clinical executive in the sample period.

@@

    Next, I plot weighted average readmission, mortality rates, and patient case mix index across time for different hospital types in Figure \ref{fig:outcomes_graph}. Pay-for-performance incentive changes are shown as a dotted line in 2012. Average readmission rates trend similarly prior to 2012, and all hospital type exhibit a decrease in readmission after 2012. However, hospitals who consistently have a clinical executive show a slower rate of decrease in raw readmissions. When considering mortality, there is not a drastic change in 2012 for any hospital type, but hospitals who always have a clinical executive continue on an increasing trend in mortality where the other hospitals types remain more flat. Finally, there do not seem to be any differential trends in case mix index prior to or after 2012.

@@

    \section{Effect of Executive Clinical Training on Hospital Response}\label{sec:clinical}

    First, I test empirically whether having clinical experience matters in how hospitals respond to the pay-for-performance policies effective in 2012. For a clear comparison, I limit the sample to hospitals who always have a clinical executive or never have a clinical executive. That is, these hospitals do not change their propensity to hire a clinical executive during the sample period. Therefore, an important assumption that I make is that this sample selection is not correlated with the program enactment. I investigate this assumption in Section \ref{sec:identification}, and find no evidence that hospitals are more or less likely to change their propensity to hire a clinical executive because of the pay-for-performance programs.

    I estimate a two way fixed effects specification, where the key variable is the interaction between having clinical executives and post 2012, when the programs began. Importantly, I am not estimating the effect of being penalized. All hospitals had incentive to improve quality as they are being compared to other hospitals who are also decreasing quality. Thus, I simply estimate the effect of being in a time period after the incentive changed for all hospitals. Ultimately, because there are differences in hospitals with different leadership types, I employ a synthetic difference-in-differences estimation strategy (\cite{arkhangelsky2021synthetic}). That is, hospitals with similar trends in readmission, mortality, and baseline case mix before the programs are more heavily weighted, and time periods that balance pre and post-program outcomes for hospitals without clinical executives. Below is the unweighted specification for clarity, where the identification assumptions are conceptually the same for the two specifications:

@@

    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:hrrp}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.

    In the unweighted two way fixed effects specification, we have reason to believe that the parallel trends assumption may not hold due to observed differences in characteristics other than the leadership team, as discussed in Section \ref{sec:data}. Once incorporating the weights used in the synthetic differences in differences method, we ensure parallel trends prior to 2012. Thus, I assume that under the given clinical and nonclinical hospital composition, trends would have continued to be parallel absent the incentive change in 2012. A component of this assumption is the implicit assumption that no other event occurred in 2012 correlated with both the outcomes and composition of leadership team. \textcolor{red}{say something about ACA here? and need to add anticipation assumption}

    The estimated difference in readmission and mortality rate trends between hospitals with and without clinical executives are shown in Figure \ref{fig:clinicalsynthdid}. Estimate $\beta$ is represented in the difference in slope after 2012 when pay-for-performance incentives began. Panel \ref{read_synth_clinical} shows the estimated difference in readmission rates, and reveals that nonclinical teams decrease readmissions by .3 more than clinical teams. This is not a huge magnitude given an average readmission rate of 21, but reveals a meaningful difference in behavior given that both hospital types significantly decrease readmissions. In panel \ref{fig:mort_synth_clinical}, I show the estimated difference in mortality. While the magnitude of the difference is similar to that for readmission rates, mortality is a noisy measure, and I cannot rule out no difference in mortality response. In Appendix \ref{app:condition}, I show that these results are primarily driven by pneumonia patients.

     \begin{figure}
     \caption{Effect of Clinical Experience on Readmission and Mortality Rates}
     \centering
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{Readmission Rate}
         \includegraphics[width=\textwidth]{Objects/read_md_nomd_synth_graph.pdf}
         \label{fig:read_synth_clinical}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{Mortality Rate}
         \includegraphics[width=\textwidth]{Objects/mort_md_nomd_synth_graph.pdf}
         \label{fig:mort_synth_clinical}
     \end{subfigure}
        \label{fig:clinicalsynthdid}
    \end{figure}

@@

    A method hospitals have been shown to employ in order to decrease readmission rates is being more selective with which patients to admit to the hospital. One hypothesis is that executive teams without any clinicians may be more willing to participate in such practices, whereas doctors prioritize care for all patients. This would explain why non-clinical hospitals are able to decrease readmissions more drastically.

    To test this, I employ case mix index as the outcome in specification \ref{eq:clinical}, to determine whether one type of hospital differentially changes patient composition after pay-for-performance incentives, where the same identification assumptions hold. I present the estimated difference in Figure \ref{fig:main_cmi_clinical}, which shows no differential response to the programs in terms of the complexity of patients admitted to the hospital. \textcolor{red}{sentence about the implications of this?}

@@

    \section{Does Specialty Matter?}

>>>>

    \begin{abstract}
		{\small
         One determinant of firm behaviors and performance is the type of management leading the firm. Characteristics such as age, gender, and background are correlated with employee wages, decisions of mergers and acquisitions, and other firm outcomes. In this paper, I investigate this topic in the context of hospital executive teams. Specifically, I estimate the effect of clinically trained executives on hospital quality. I leverage pay-for-performance policies that changed quality incentives for hospitals in the US in 2012, and estimate the difference in readmission and mortality rates for nonprofit hospitals with and without clinically trained executives. I find that clinically trained executives do not improve quality as much as executive teams without clinically trained executives. This difference is not driven by a change in patient complexity, but may be driven by preferences towards profit vs. patient welfare. Further, clinically trained executives are not just a signal of underlying hospital incentives, but manage the hospital differently. 
		} 
	\end{abstract}

@@

    Broadly, firm leader characteristics such as gender, age, and background have been shown to be correlated with financial performance and behaviors such as employee diversity, wages, aggressiveness in mergers and acquisitions, and other corporate strategies. I consider this topic in the context of hospital quality under executives with differing backgrounds: some with training as a medical doctor and some without. Physicians can bring a unique combination of clinical and administration expertise, and they have insights that can potentially benefit their employees and lead to better quality (\cite{Stajduhar_2023}, \cite{Ahmed_2022}). However, hiring physician executives can also be controversial, as physicians are not always adequately trained in business management, which can lead to poor business practices and lower quality (\cite{HarvardBusinessReview2018}).

    In this paper, I estimate how having clinically trained executives in leadership affects quality of care in hospitals. I scrape executive team names and titles from publicly available Tax Form 990s, filled out by nonprofit firms in the US. Combining this with publicly available hospital characteristics, I construct a novel data set of nonprofit hospitals from 2010-2014 containing information on both executive team characteristics and quality metrics. While limited to nonprofit hospitals due to data availability, this sample is a very important driver of health care in the US. Private nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds. For comparison, for-profits make up 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}). Therefore, the behavior of these hospitals directly affects health care consumers in the US.

    The non-random selection of leaders makes it difficult to interpret any direct comparison of differing executive teams as causal. Therefore, I leverage an exogenous shock to hospital incentives, pay-for-performance policies enacted in the US in 2012, and compare the change in readmission and mortality rates of executive teams with and without clinical executives. Under the assumption that the two types of teams would have had similar trends in readmission and mortality rates absent the incentive change, I find that executive teams with clinical executives on average do not improve quality as much as executive teams without clinical executives. While both types of hospital leadership teams decrease readmission rates after the policy change, those without clinical executives decrease readmissions at a faster rate than those with clinical executives.

    Having found that hospital leadership does, in fact, matter to hospital behavior, I investigate several mechanisms that might contribute to this difference. First, the more drastic decrease in readmissions by non-clinical teams is not driven by decreasing average patient complexity. Further, comparing each type of leadership team with for-profit hospitals reveals that the absence of clinical executives leads a nonprofit hospital to respond more similarly to a for-profit hospital than nonprofits with clinical executives. That is, having clinical training on a leadership team leads a hospital to act less profit-driven than their general training counterparts. Finally, I show that changes in a hospital's propensity to hire a clinical executive are not correlated with the pay-for-performance policies. Thus, I exploit such changes in leadership teams to decompose the effect of clinical executives into signaling (clinical executives signal underlying hospital preferences), or managing (clinical executives manage the hospital differently). I find that the entire difference in response is driven by a managing effect.

    Many researchers have documented a correlation between executive characteristics and firm performance, starting with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who show that manager fixed effects are an important driver of many firm behaviors and decisions. Female board members are associated with better oversight and more women executives (\cite{matsa2011chipping}; \cite{adams2009women}), but are negatively correlated with firm value (\cite{ahern2012changing}). Having more female executives is correlated with female employee wages and corporate strategies (\cite{flabbi2019female}; \cite{matsa2013female}); young male CEOs tend to be more aggressive in mergers and acquisitions, while those with military experience are less aggressive (\cite{levi2010deal}; \cite{benmelech2015military}); CEOs with general ability tend to receive higher pay and perform better (\cite{kaplan2012ceo}; \cite{custodio2013generalists}; \cite{adams2018director}; \cite{frydman2019rising}). Finally, Chief Diversity Officers are not found to have any effect on hiring more diversely in universities (\cite{bradley2022impact}).

    Due to data limitations, many of these studies focus on large, publicly traded firms in the US. \citeauthor{brickley2010board} (\citeyear{brickley2010board}) have the only study to my knowledge in the US nonprofit context considering hospital board of directors. Using exogenous variation in expected Medicare profits, the authors find that having an internal board of directors increases CEO compensation, and having physician board members decreases public donations (\cite{brickley2010board}). Additionally, two studies focus on hospital performance in other countries. \citeauthor{janke2019impact} (\citeyear{janke2019impact}) use data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}). However, \citeauthor{otero2022managers} (\citeyear{otero2022managers}) investigate the role of CEOs in public hospitals in Chile, and finds an 8\% decrease in mortality rates for hospitals with top managers. Further, a major contributor of this quality improvement is the exodus of older physicians as CEOs (\cite{otero2022managers}). I contribute to our knowledge of how executive team characteristics affect firm behaviors by (1) considering the important setting of nonprofit hospitals, which serve the majority of patients in the US, (2) considering a unique and relevant executive characteristic, and (3) observing firm behavior in the presence of an exogenous incentive change.

@@

    Nonprofit firms in the US differ from for-profits mainly in that for-profit firms distribute revenue to stakeholders, while nonprofits reinvest profit back into the firm. Hospitals are governed by a board of directors, whose role is to set broad goals and strategies, and provide general oversight. The board selects executives, the highest level of management of a firm, to carry out day-to-day operations. An executive team usually consists of at least a Chief Executive Officer (CEO) and Chief Financial Officer (CFO), but there is variation in how firms organize executive teams. Some hospital executives specialize in health care settings by earning a degree in health care management, or an MBA specific to health care. There are executive positions that are often filled by someone with clinical training, such as a Chief Medical Officer (CMO), but medical doctors can pursue other top managerial positions as well such as CEO or president. While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.

    Hospital executive teams are understudied partly due to the inaccessibility of granular data. I construct a novel data set of nonprofit hospitals in the US which contains names, titles, and positions of all board members and executives tied to the hospital from 2009-2015. I gather this data from publicly available Tax Form 990s, which every sufficiently large nonprofit files each year in the US. To my knowledge, this is the first large scale gathering of executive names from these forms.\footnote{\citeauthor{brickley2010board} (\citeyear{brickley2010board}) collects compensation data for a select number of hospitals.}

@@

    The historical tax forms are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the NonProfit Explorer API to extract Employee Identification Numbers (EIN) of firms with a hospital National Taxonomy of Exempt Entities (NTEE) code. After filtering out associations and specialty hospitals, there are approximately 3,000 EINs. Linking these EINs to other hospital information is crucial to the analysis, but no crosswalk from EIN to other hospital identifiers is publicly available. I therefore rely on matching by location and name of hospital in both the tax forms and American Hospital Association (AHA) data. In the AHA data, I limit to general hospitals. Then, I confidently match 1200 EINs to an AHA identification number based on exact name matches within the same state.\footnote{I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.}

    For each of the matched EINs, I extract web URLs to Form 990 PDFs for each year. I use an algorithm to download these locally and use optical character recognition (OCR) text extraction methods to scrape the relevant section of the PDF, which is titled ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees". Finally, I use string cleaning methods to identify all names and positions under each EIN, year. With each name, I extract titles indicating clinical training: MD, Dr., or DO. Using OCR extraction can be unreliable in cases where the text on the form is handwritten or small. Therefore, for observations that are missing after the initial extraction, I manually record names, positions, and titles. As I limit the sample to hospitals who have text information in every year, I limit the years to 2010-2014 to retain more hospitals. This yields 852 nonprofit hospitals that are matched to AHA data, and with complete text information in each year.

    Finally, I seek to (1) verify clinical training of those who claim to be doctors, and (2) obtain additional information about clinically trained executives. Thus, I match each name with the National Plan and Provider Enumeration System (NPPES) database of all registered physicians. For any person who has a doctor title and uniquely matches the NPPES data, I record their National Provider Identifier (NPI) number. For anyone who has a doctor title and no matches, no doctor title and matches, or a doctor title and multiple matches, I use information on location and hospital name to manually match NPI numbers of those who did have clinical training. I present a table of means for various characteristics of doctor executives in the data in Table \ref{doc_sumstats}. In total, there are 813 doctor executives. On average, they are 52 years old, and 11\% of them are female. While it is more likely they are a Chief Medical Officer than a Chief Executive Officer, still less than a third are CMOs. The most common specialty is internal medicine.

@@

    For the purposes of this paper, I focus on executive teams as a whole instead of specific positions. The main reason for this is because the organization of teams varies so drastically across hospitals. It is not uncommon for two hospitals to have the same position but the tasks performed are drastically different, or to have two distinct positions that essentially perform the same tasks. I drop anyone who is solely a board member, and focus on those who have executive in their title, or are labeled as a president or vice president of a department. I create various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer. For the main analysis, I limit to hospitals who either always have a clinical executive or never have a clinical executive over the course of the sample in order to have a clean comparison.

@@

    I focus on outcomes of readmission and mortality rates, which are publicly available at the hospital-condition level in CMS Hospital Compare. I combine rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions) into a weighted average based on the number of patients in each condition. Additionally, I include several relevant hospital characteristics from the AHA survey: number of beds, and whether the hospital is an academic medical center, is physician owned, or is system affiliated. Additionally, case mix index, a measure of overall patient complexity, comes from the CMS Case Mix Index files. Whether the hospital was penalized under HRRP or received payments under HVBP is found in the HCRIS data.

@@

    The hospitals in my sample are limited to general nonprofit hospitals with executive team information in each year (2010-2014) and a link from EIN to AHA ID. In Table \ref{tab:sumstats_samples_stable}, I present means of relevant variables for several hospital samples used throughout the paper. First, in column (1), are means for hospitals who ever have a clinical executive at any point in the sample. In column (2) are means for hospitals who have a clinical executive for the entire sample period, and column (3) shows means for hospitals who never have a clinical executive in the sample period. The main analysis compares hospitals in columns (2) and (3), and supplemental analyses leverage hospitals in column (1).

@@

    Next, I plot weighted average readmission, mortality rates, and patient case mix index across time for different hospital types in Figure \ref{fig:outcomes_graph}. Pay-for-performance incentive change timing is shown as a dotted line in 2012. Average readmission rates trend similarly prior to 2012, and all hospital type exhibit a decrease in readmission after 2012. However, hospitals who consistently have a clinical executive show a slower rate of decrease in raw readmissions. When considering mortality, there is not a drastic change in 2012 for any hospital type, but hospitals who always have a clinical executive continue on an increasing trend in mortality where the other hospitals types remain more flat. Finally, hospitals that never have clinical executives consistently have less complex patients, but there is no notable change in raw case mix index over time.

@@

    \section{Effect of Executive Clinical Training on Hospital Quality}\label{sec:clinical}

    First, I test empirically whether having clinical training matters in how hospitals change readmission and mortality rates in response to the pay-for-performance policies effective in 2012. For a clean comparison, I limit the sample to hospitals who always have clinical executives or never have clinical executives. That is, these hospitals do not change their propensity to hire a clinical executive during the sample period. Therefore, an important assumption that I make is that this sample selection is not correlated with the program enactment. I investigate this assumption in Section \ref{sec:changes}, and find no evidence that hospitals are more or less likely to change their propensity to hire a clinical executive because of the pay-for-performance programs.

    I estimate a two-way-fixed-effects specification, where the key independent variable is the interaction between having clinically trained executives and post 2012, when the programs began. Importantly, I am not estimating the effect of being penalized. All hospitals had incentive to improve quality in response to the policy change, as penalties and payments are determined by comparing quality among hospitals. Thus, I leverage the timing of the policy change, rather than the policies themselves. However, because of potential differences in quality trends between clinical and non-clinical hospital executive teams even before the policies, I employ a synthetic difference-in-differences estimation strategy (\cite{arkhangelsky2021synthetic}). That is, hospitals in the control group with similar trends in readmission, mortality, and baseline case mix before the programs are more heavily weighted, as well as time periods that balance pre- and post-program outcomes for hospitals without clinical executives. Below is the unweighted specification for clarity, where the identification assumptions are conceptually the same for synthetic difference-in-differences:

@@

    \noindent where $y_{ht}$ is one of the outcome variables capturing readmission or mortality rates discussed in Section \ref{sec:hrrp}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.

    In the unweighted two-way-fixed-effects specification, we have reason to believe that the parallel trends assumption may not hold due to observed differences in characteristics other than the leadership team, as discussed in Section \ref{sec:data}. Once incorporating the weights used in the synthetic differences in differences method, we ensure parallel trends in the outcome variables prior to 2012. Thus, I assume that under the weighted clinical and non-clinical hospital composition, trends would have continued to be parallel absent the incentive change in 2012. A component of this assumption is the implicit assumption that no other event occurred in 2012 correlated with both the outcomes and composition of leadership team. Finally, I assume that there is no anticipation of the policy change before 2012. The rules of the policies were announced in October of 2011, so 2012 was the first full year that hospitals could meaningfully respond to the policies.

    The estimated difference in readmission and mortality rates between hospitals with and without clinical executives are shown in Figure \ref{fig:clinicalsynthdid}. Estimate $\beta$ is represented in the difference in slope after 2012 when pay-for-performance incentives began. Panel \ref{fig:read_synth_clinical} shows the estimated difference in readmission rates, and reveals that non-clinical teams decrease readmissions by .3 (standard error .13) more than clinical teams. This is a relatively small magnitude given an average readmission rate of 21, but reveals a meaningful difference in behavior given that both hospital types significantly decrease readmissions. In panel \ref{fig:mort_synth_clinical}, I show the estimated difference in mortality. While the magnitude of the difference is similar to that for readmission rates, mortality is a noisy measure, and I cannot rule out no difference in mortality response. In Appendix \ref{app:condition}, I show that these results are primarily driven by pneumonia patients.

     \begin{figure}[ht!]
     \caption{Effect of Clinical Experience on Readmission and Mortality Rates}
     \centering
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{Readmission Rate}
         \includegraphics[width=\textwidth]{Objects/read_md_nomd_synth_graph.pdf}
         \label{fig:read_synth_clinical}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{Mortality Rate}
         \includegraphics[width=\textwidth]{Objects/mort_md_nomd_synth_graph.pdf}
         \label{fig:mort_synth_clinical}
     \end{subfigure}
        \label{fig:clinicalsynthdid}
    \end{figure}

    Additionally, I estimate the intensive margin of clinical leadership: whether the fraction of clinical executives on the team affects quality. I bin the fraction of clinical executives into less than a quarter or more than a quarter and less than one half.\footnote{Less than .5\% of hospitals have greater than 50\% clinical executives on the team.} Then, I estimate:

    \begin{equation}
    \label{eq:clinical_continuous_onefourth}
    y_{ht} = \beta \left(\mathbbm{1}\{0 < \rho \leq .25\}\text{ x post 2012}\right)_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}, \text{ and}
    \end{equation}
    \begin{equation}
    \label{eq:clinical_continuous_onehalf}
    y_{ht} = \beta \left(\mathbbm{1}\{.25 < \rho \leq .5\}\text{ x post 2012}\right)_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht},
    \end{equation}

    \noindent where $\rho$ is the fraction of clinical executives on hospital $h$'s executive team, and $\gamma$ and $\delta$ are fixed effects as defined in Equation \ref{eq:clinical}.

    I show the results of this estimation with readmission rates as the outcome in Figure \ref{fig:cont_clinicalsynthdid}. In panel \ref{fig:onefourth_read_synth_clinical}, the estimation compares executive teams with no clinical experience to those with up to 25\% of the team having clinical experience. The average treatment effect is .04 and not statistically significant. Thus, there is no significant difference in how these types of teams respond to the incentive change. In panel \ref{fig:onehalf_read_synth_clinical}, I compare hospitals with no clinically trained executives over the course of the sample to hospitals with over a quarter of their team having clinical training. While these hospitals have parallel pre-trends, hospitals without clinically trained executives decrease readmissions by .64 (standard error .33) more than hospitals with clinically trained executives. Thus, the readmission rate response is fully driven by having a higher proportion of clinical executives relative to no clinical executives.

    \begin{figure}[ht!]
     \caption{Effect of Clinical Experience on Readmission Rates, Binned Treatment}
     \centering
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{$0 < \rho \leq .25$}
         \includegraphics[width=\textwidth]{Objects/cont_onefourthread_md_nomd_synth_graph.pdf}
         \label{fig:onefourth_read_synth_clinical}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{$.25 < \rho \leq .5$}
         \includegraphics[width=\textwidth]{Objects/cont_onehalfread_md_nomd_synth_graph.pdf}
         \label{fig:onehalf_read_synth_clinical}
     \end{subfigure}
        \label{fig:cont_clinicalsynthdid}
    \end{figure}

    Any potential difference in mortality rate response also seems to be driven by executive teams with over a quarter of the executives having clinical experience, as shown in Figure \ref{fig:cont_clinicalsynthdid_mort}. However, the standard errors of this estimation are high, potentially due to the noisiness in the mortality measure. The magnitude of the estimate in panel $\ref{fig:onefourth_mort_synth_clinical}$ is .07, so I can rule out differential responses by teams with less than a quarter of clinical executives. Taking the readmission and mortality rate results together, the effect of having clinical leadership on quality after an incentive change depends on having a sufficient proportion of the team with clinical training.

    \begin{figure}[ht!]
     \caption{Effect of Clinical Experience on Mortality Rates, Binned Treatment}
     \centering
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{$0 < \rho \leq .25$}
         \includegraphics[width=\textwidth]{Objects/cont_onefourthmort_md_nomd_synth_graph.pdf}
         \label{fig:onefourth_mort_synth_clinical}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{$.25 < \rho \leq .5$}
         \includegraphics[width=\textwidth]{Objects/cont_onehalfmort_md_nomd_synth_graph.pdf}
         \label{fig:onehalf_mort_synth_clinical}
     \end{subfigure}
        \label{fig:cont_clinicalsynthdid_mort}
    \end{figure}

@@

    A method hospitals have been shown to employ in order to decrease readmission rates is being more selective with which patients to admit to the hospital (\cite{gupta2021impacts}). One hypothesis behind a more drastic quality decrease by non-clinical teams is that they may be more willing to participate in such practices, while executives with clinical training prioritize care for all types of patients.

    To test this, I employ case mix index as the outcome in specification \ref{eq:clinical} to determine whether one type of hospital differentially changes patient composition after pay-for-performance incentives, where the same identification assumptions hold. I present the estimated difference in Figure \ref{fig:main_cmi_clinical}, comparing hospitals that never have a clinical executive to hospitals that always have a clinical executive. Neither type of hospital drastically changes case mix index after the incentive change, and there is no estimated differential response in complexity of patients seen in the hospital. Thus, the estimated difference in readmission rate response between clinical and non-clinical executives is not due to differentially changing case mix.

@@

    \section{Internal Medicine vs. Other Specialties}
